---
title: "DL_Data_Collection_Intern"
output: html_document
date: '2022-08-15'
---

Loading in packages used for this program
```{r}
install.packages('here')
install.packages('tidyverse')
install.packages('sf')
install.packages('janitor')
install.packages('dplyr')
install.packages('rgdal')
library(janitor)
library(dplyr)
library(tidyverse)
library(sf)
library(here)
library(rgdal)
library(sp)
```


Reading in shapefiles using st_read from sf package.
```{r}

graz_sf <- here("/Users/daniellopez/Downloads/Internship Data Packs/Projects_Grazing.shp") %>%
  st_read()
veg_comp_sf <- here("/Users/daniellopez/Downloads/Internship Data Packs/SB_City_Completed_Veg_Projects.shp") %>%
  st_read()
mixed_sf <- here("/Users/daniellopez/Downloads/Internship Data Packs/Projects_MixedTreatment.shp") %>%
  st_read()
fuel_treatments_sf <- here("/Users/daniellopez/Downloads/Internship Data Packs/Fuel_Treatment_Polygons_20220523.shp") %>%
  st_read()
sbc_future_sf <- here("/Users/daniellopez/Downloads/Internship Data Packs/SB_City_Future_Projects.shp") %>%
  st_read()

```


Viewing coordinate reference system of shapefiles
```{r}

st_crs(graz_sf)
st_crs(veg_comp_sf)
st_crs(mixed_sf)
st_crs(fuel_treatments_sf)
st_crs(sbc_future_sf)

```


Re-projecting crs(coordinate reference system) of two shapefiles to match other 3 as [NAD83 CRS] from [WGS 84]
```{r}

graz_sf <- st_transform(graz_sf, st_crs(fuel_treatments_sf))
st_crs(graz_sf)

mixed_sf <- st_transform(mixed_sf, st_crs(fuel_treatments_sf))
st_crs(mixed_sf)

```


Selecting the columns only wanted from shapefile attribute tables
```{r}

graz_sf <- select(graz_sf, "Original Shapefile Name", "Agency", "Date Acquired","Acres", "Treatment_", "Timeframe_", "Date Completed", "Other")

mixed_sf <- select(mixed_sf, "Original Shapefile Name", "Agency", "Date Acquired", "Project_Na", "Acres", "Treatment_", "Other")

veg_comp_sf <- select(veg_comp_sf, "SHAPE_Area", "TREATMENT_", "WorkArea", "FUEL_MGMT", "Treatment")

fuel_treatments_sf <- select(fuel_treatments_sf, "project_na", "acres", "poly_type")

sbc_future_sf <- select(sbc_future_sf, "VMU_Name", "Shape_STAr")

```


Adding columns wanted by using mutate + the shapefile name.
```{r}

##
graz_sf <- mutate(graz_sf, 
                  "Agency" = "Montecito", 
                  "Project Name" = "NA",
                  "Date Acquired" = "Jul 11, 2022",
                  "Date Completed" = "Proposed Treatments",
                  "Original Shapefile Name" = "Projects_Grazing", 
                  "Other" = "NA"
)
##

##
mixed_sf <- mutate(mixed_sf, 
                   "Agency" = "Montecito", 
                   "Date Acquired" = "Jul 11, 2022", 
                   "Date Completed" = "Proposed Treatments",
                   "Original Shapefile Name" = "Projects_MixedTreatment", 
                   "Other" = "NA"
)
##

##
veg_comp_sf <- mutate(veg_comp_sf, 
                    "Agency" = "Santa Barbara City",
                    "Date Acquired" = "Jul 12, 2022", 
                    "Original Shapefile Name" = "SB_City_Completed_Veg_Projects"
)
##

```


```{r}

##
sbc_future_sf <- mutate(sbc_future_sf, 
                    "Agency" = "Santa Barbara City",
                    "Date Acquired" = "Aug 2, 2022",
                    "Date Completed" = "NA",
                    "Treatment Type" = "NA",
                    "Original Shapefile Name" = "SB_City_Future_Projects", 
                    "Other" = "NA"
)

##

##

fuel_treatments_sf <- mutate(fuel_treatments_sf, 
                             "Agency" = "Santa Barbara County",
                             "Date Acquired" = "Jul 6, 2022",
                             "Date Completed" = "NA",
                             "Original Shapefile Name" = "Fuel_Treatment_Polygons_20220523", 
                             "Other" = "NA"
)

##

```


Renaming columns for grazing, mixed fuel treatments, vegetation completed, fuel treatments, & SB city completed project shapefiles.
```{r}
##
graz_sf <- rename(graz_sf, 
                  "Treatment Type" = "Treatment_",
                  "Time Frame (Days)" = "Timeframe_",
                  "Acreage" = "Acres"
)
##

##
mixed_sf <- rename(mixed_sf, 
                   "Project Name" = "Project_Na",
                   "Treatment Type" = "Treatment_",
                   "Acreage" = "Acres"
)
##

##
veg_comp_sf <- rename(veg_comp_sf, 
                      "Treatment Type" = "TREATMENT_", 
                      "Other" = "Treatment", 
                      "Project Name" = "WorkArea", 
                      "Date Completed" = "FUEL_MGMT", 
                      "Acreage" = "SHAPE_Area"
                      
)
##

##
fuel_treatments_sf <- rename(fuel_treatments_sf, 
                             "Project Name" = "project_na", 
                             "Acreage" = "acres",
                             "Treatment Type" = "poly_type"
)
##

##
sbc_future_sf <- rename(sbc_future_sf, 
                        "Project Name" = "VMU_Name", 
                        "Acreage" = "Shape_STAr"
)

```


Merging together data tables and shapefiles
```{r}

merge_1 <- bind_rows(graz_sf, 
                     mixed_sf
)

"This below I used to reorder the columns how I wanted them"
merge_1 <- select(merge_1, 
          1, 2, 3, 
          9, 4, 5, 
          6, 7, 8
)


##

merge_2 <- bind_rows(graz_sf, 
                     mixed_sf, 
                     veg_comp_sf
)

merge_2 <- select(merge_2, 
          1, 2, 3, 
          9, 4, 5, 
          6, 7, 8
)

##

##
merge_3 <- bind_rows(graz_sf, 
                     mixed_sf, 
                     veg_comp_sf, 
                     sbc_future_sf
)

merge_3 <- select(merge_3, 
           1, 2, 3, 
          9, 4, 5, 
          6, 7, 8
)

##

##

merge_4 <- bind_rows(graz_sf,
                     mixed_sf,
                     veg_comp_sf,
                     sbc_future_sf,
                     fuel_treatments_sf
)

##

merge_4 <- select(merge_4, 
          1, 2, 3, 
          4, 6, 9, 
          5, 7, 8
)

##

```


Exporting full data merge of 5 shapefiles as a csv file and as a shapefile
```{r}

st_write(merge_4, "Complete Data Merge.csv")
st_write(merge_4, "Complete_Data_Merge.shp")

```


This is reading in the Carpinteria shapfile.
```{r}

carp_sf <- here("/Users/daniellopez/Downloads/Internship Data Packs/Carpinteria Merge.shp") %>%
  st_read()

```


By using "select" I am choosing which columns I want from the carpinteria shapefile data table.
```{r}

carp_sf <- select(carp_sf, "Activity", "fuel_type", "Acres", "Name_Treat", "Comp_Date","layer")

```


Renaming the columns for the carpinteria shapefile below.
```{r}

carp_sf <- rename(carp_sf, 
                  "Acreage" = "Acres",
                  "Treatment Type" = "Activity",
                  "Project Name" = "Name_Treat",
                  "Other" = "fuel_type",
                  "Date Completed" = "Comp_Date",
                  "Original Shapefile Name" = "layer"
)

```


Adding 2 extra  needed columns for the Carpinteria data
```{r}

carp_sf <- mutate(carp_sf, 
                  "Date Acquired" = "Aug 23, 2022",
                  "Agency" = "Carpinteria"
)

```


Below is needed for merge to work as "Date Completed" for the carp data is read in as a "double" and not "character" like the other 5 shapefile data so by using as.character I set it from double to character so merge can be completed without error.
```{r}

typeof(carp_sf$`Date Completed`)

carp_sf$`Date Completed` <- as.character(carp_sf$`Date Completed`)

```


Below [merge_5] is only required if needed but I am merging 6 of the current shapfiles together then using st_write to export a single merged shapefile for GIS software.
```{r}

merge_5 <- bind_rows(graz_sf,
                     mixed_sf,
                     sbc_future_sf,
                     veg_comp_sf,
                     fuel_treatments_sf, 
                     carp_sf
)

st_write(merge_5, "merge_5.shp")

```


Below I am reading in the new/additional shapefile
```{r}

rob_sbc_sf <- here("/Users/daniellopez/Downloads/Internship Data Packs/ROBH_DATA.shp") %>%
  st_read()

```


After reading in the shapefile from above I am using select to pick the columns I only want and will be using from the shapefile
```{r}

rob_sbc_sf <- select(rob_sbc_sf, 
                     "Acres", 
                     "Name", 
                     "Agency", 
                     "Fuel_Type", 
                     "Activity", 
                     "layer",
)

```


Here I am renaming the columns to the requested names
```{r}

rob_sbc_sf <- rename(rob_sbc_sf, 
                     "Acreage" = "Acres", 
                     "Project Name" = "Name", 
                     "Treatment Type" = "Activity", 
                     "Original Shapefile Name" = "layer", 
                     "Other" = "Fuel_Type"
)

```


I am adding 2 extra columns needed for data table using "mutate" 
```{r}

rob_sbc_sf <- mutate(rob_sbc_sf, 
                     "Date Acquired" = "Aug 29, 2022", 
                     "Date Completed" = "NA"
)


```


Below I am merging all shape files together and then using "select" to reorder the columns to preference
```{r}

merge_6 <- bind_rows(graz_sf,
                     mixed_sf,
                     sbc_future_sf,
                     veg_comp_sf,
                     fuel_treatments_sf,
                     carp_sf,
                     rob_sbc_sf
)


merge_6 <- select(merge_6, 
          1, 2, 3, 
          4, 6, 9, 
          5, 7, 8
)

"st_write is me exporting the final merge so that it can be opened in GIS software as a shapefile"

st_write(merge_6, "Final_Merge.shp")
st_write(merge_6, "Final_Merge.csv")

```


~ By: Daniel Lopez
September 16, 2022
